---
title: "Project02"
author: "Max Selvaggio"
date: "2022-07-31"
output: html_document
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(pROC)
library(rpart)
library(rpart.plot)
library(gamlr)
```


```{r}
setwd("C:\\Users\\mselv\\Documents\\GitHub\\ECO5445\\Project\\Data")
Mdata <- read.csv("hmda_sw.csv")

setnames(Mdata, old = c('s13','s46','s27a','s23a','school','s17','s7'), 
         new = c('Applicant Race','Debt to Income Ratio','Self-Employed','Marital Status',
                 'Years of Education','Applicant Income','Action Taken'))

Mdata[Mdata == 999999.4] <- NA
Mdata$`Action Taken`[Mdata$`Action Taken` == 2] <- 1
Mdata$`Action Taken` <- factor(Mdata$`Action Taken`,levels = c(1,3), labels = c("Approved","Denied"))

Mdata["Applicant Race"][Mdata["Applicant Race"] == "3"] <- "Black"
Mdata["Applicant Race"][Mdata["Applicant Race"] == "5"] <- "White"

Mdata["Self-Employed"][Mdata["Self-Employed"] == "0"] <- "No"
Mdata["Self-Employed"][Mdata["Self-Employed"] == "1"] <- "Yes"

Mdata$`Applicant Income` <- Mdata$`Applicant Income`*1000
```

The variable "Action Taken" is our qualitative dependent because the approval or denial of a mortgage is the output we are attempting to predict with this model. The reason these variables were chosen is because it is important to have "variables correlated with both race's role and creditworthiness"(Page 1, "Mortgage Lending in Boston: Interpreting HMDA Data"). As the paper indicates, previous studies of how race played a part in the mortgage process did not properly include financial variables as well, and therefore the results were skewed. With these variables, we will be able to see race's affect as well as the affect of general financial measures in the mortgage approval process.

```{r}
newdata <- Mdata[,c(6,9,13,20,24,40,59)]

summary(newdata)


table1 <- table(newdata$`Applicant Race`) 
table2 <- table(newdata$`Self-Employed`)
table3 <- table(newdata$`Marital Status`)

hist(newdata$`Applicant Income`, xlim = c(1,1000000), breaks = 20)
hist(newdata$`Debt to Income Ratio`, xlim = c(1,100), breaks = 30)




```

Notes from different graphs and summary statistics:

Based off "table1", you see in the data set 2041/2380 (86%) of applicants are white. "table2" demonstrates that the majority of applicants are not self-employed, 2103/2380 (88%). The histogram of the Applicants' annual income shows that most of these people fall into a similar ranges of income. 812 applicants fall into the 0-50,000 range, and 1130 applicants fall into the 50,001-100,000 range. It is somewhat intuitive, but it still important to keep in mind that the average applicant is making under 100,000 annually because if the ranges were more evenly spread out across the histogram than the approval prediction would be heavily concerned with the range in which an applicant falls. Instead, the applicants are more or less on an even playing field except for the few outliers making upwards of 500,000. Finally, the histogram of the applicants' Debt to Income Ratio shows the majority of people fall into the 20 to 40 range, which are good ranges in the eyes of the bank as the recognized ideal range is around 28-36. Taking these charts and tables into account, you can see that the average applicants is married, white, employed, making under 100,000 annually and has a solid debt to income ratio in the range of 20 to 40 percent.

```{r}
set.seed(1234)

train <- sample(nrow(newdata), 0.7*nrow(newdata))

newdata.train <- newdata[train,]
newdata.validate <- newdata[-train,]

table(newdata.train$`Action Taken`)
summary(newdata.validate$`Action Taken`)

model <- glm(`Action Taken` ~., data = newdata.train, family = "binomial")

summary(model)

```


```{r}
prob <- predict(model, newdata.validate, type = 'response')

model.pred <- factor(prob > 0.5, levels = c(FALSE,TRUE),
                     labels = c("Approved","Denied"))

model.perf <- table(newdata.validate$`Action Taken`, model.pred, dnn = c("Actual","Predicted"))

model.perf

my_roc <- roc(newdata.validate$`Action Taken` ~ prob, plot = TRUE)

my_auc <- auc(my_roc)


```


```{r}
threshold <- my_roc$thresholds[which.max(my_roc$sensitivities + my_roc$specificities)]

model.pred.thresh <- factor(prob > threshold, levels = c(FALSE,TRUE),
                     labels = c("Approved","Denied"))

model.perf.thresh <- table(newdata.validate$`Action Taken`, model.pred.thresh, 
                     dnn = c("Actual","Predicted"))

model.perf.thresh


```


When the threshold was put at 0.5, the model produced a really good number of true approvals, but also yielded 75 false positives (approvals). The threshold taken from the ROC curve was used and produced many more false negatives (predicting denied when actually approved), but limited the false positives (predicting approved when actually denied), reducing them from 75 to 42. Similar to our breast cancer example from class, I think we would prefer the higher number of false negatives rather than false positives, which give applicants unwarranted hope.


```{r}

x <- data.matrix(newdata.train[, c('Applicant Race', 'Applicant Income', 'Marital Status', 'Self-Employed', 
                                  'Debt to Income Ratio', 'Years of Education')])

y <- newdata.train$`Action Taken`

cv.model <- cv.gamlr(x, y, 
	family="binomial", nfold = 10, verb=TRUE)
```


