---
title: "Assignment05"
author: "Max Selvaggio"
date: '2022-07-16'
output: html_document
---
*JE: Would have liked to have had same YAML header with table of contents*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*JE: Removed unnecessary content in first page*

# Number 2 Assignment 2 

*JE: Adding `#` so that the sections are listed properly*

```{r}
Raft <- 2
Staff <- 2.0
Nice <- "10j" # Did you try without the quotes? 
Goat <- "2 Cool for School"
Jump <- "True" # Did you try without the quotes?


class(Raft) # Interesting, didn't realize there was a typeof function
typeof(Staff)
typeof(Nice)
class(Goat)
typeof(Jump)
# Fixed weird indent
```

# Number 5 Assignment 2

foobar is a function which determines if a numerical value is a multiple of 3,5, or 15.

*JE: Did not provide examples, did not run any tests, errors out if character is provided in function. (-6)*

```{r}
foobar<-function(x){
    return(ifelse(x%%3==0 & x%%5==0, "foobar", ifelse(x%%3==0, "foo", ifelse(x%%5==0, "bar", "Not a multiple of 3, 5, or 15"))))
}

# foobar("cat") commented out to allow document run
foobar(90)
foobar(20)
foobar(9)
```

# Number 3

```{r}
attach(swiss)
```

# Number 4

```{r}
summary(swiss)


```

*JE: Didn't provide spread of data (sd or variance) (-5)*

# Number 5  -  Herens

*JE: Where is the code to show that? (-10)*

# Number 6

```{r}
cor(swiss)
```

# Number 7

```{r}
plot(Education,Fertility)
```

# Number 8

```{r}
hist(Catholic)
```

```{r}
detach(swiss)
```

